---
title: "Final Project Summary"
author: "Lauren Berger"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
  word_document: default
bibliography: GRFP_new.bib
csl: apa.csl
---


**Historical contingency in post-fire plant communities: analyzing diversity in managed forests**

**Author:** Lauren Berger 


## I. Introduction

Wildfires in the western U.S. have become more frequent and severe[@halofskyChangingWildfireChanging2020a]. High-severity
fires (>95% tree mortality) have increased eight-fold due to long-term fire exclusion and climate change[@halofskyChangingWildfireChanging2020a, @parksWarmerDrierFire2020]. Meanwhile, many forests in this region are managed for timber production. As a result, the compounded disturbances of high-severity fire and logging increasingly co-occur, potentially diminishing biodiversity beyond either event alone. While mixed-severity fires create habitat mosaics that support biodiversity, [@ponisioPyrodiversityPromotesInteraction2020a] high-severity fires often simplify communities by reducing local diversity and species turnover[@steelWhenBiggerIsnt2022]. However, few studies have examined the influence of historical contingency in driving these patterns. With wildfire severity and prevalence projected to rise, [@parksWarmerDrierFire2020] understanding how fire shapes community assembly in managed forests is crucial for predicting and managing biodiversity at local and regional scales. 

Historical contingency provides a framework to interpret the effects of varying disturbance legacies on community assembly, leading to measurable differences between sites under similar conditions[@fukamiHistoricalContingencyCommunity2015]. Two crucial components of historical contingency —site legacies (enduring characteristics from past events) [@johnstoneChangingDisturbanceRegimes2016] and priority effects (sequence and timing of species arrivals) [@fukamiHistoricalContingencyCommunity2015]—are particularly relevant when assessing post-fire ecosystems.

The only two studies on site legacies in harvested forests found that past stand age influences plant composition. Older stands tend to regenerate shrub-dominated assemblages resembling pre-fire communities, while younger stands often contain more disturbance-adapted vegetation [@tonEffectsDisturbanceHistory2016], [@bowdPriorDisturbanceLegacy2021]. Yet, effects on local species ($\alpha$ diversity) and turnover ($\beta$ diversity) remain under-examined. Furthermore, research on plant response to fire rarely considers the temporal order of community assembly. As high-severity fire often eliminates existing propagules, input from nearby refugia is necessary for re-colonization.[@cottrellSeedInvasionFilters2008], [@ulyshenForestPollinatorRichness2024]. Given the increasing extent of high severity fire, priority effects may intensify filtering, favoring plants with traits capable of withstanding high severity fire or greater dispersal. While these early-arriving species can rapidly establish populations, they often overlap in ecological roles, leading to communities with higher functional redundancy but lower diversity.

Due to the influence of historical contingency, I predict that variation in biological legacies associated with pre-fire stand age and subsequent species filtering via priority effects will generate measurable differences in post-fire plant communities. Specifically, I predict that sites embedded within increasingly high-severity patch areas will exhibit: (H1) lower taxonomic plant $\alpha$ diversity and (H2) lower $\beta$ diversity. In all hypotheses, I expect stand age pre-fire to moderate these effects: older stands retain biological legacies that buffer diversity loss, while younger stands display exaggerated responses.

Note to reader: This introduction was originally written with the goal of including stand age in my analysis. While the code was completed to incorporate stand age, ultimately the data was inaccurate. I have kept mention of stand age in my introduction as it will be included in future analysis. The code for stand age is included at the end of this RMD.

## II. Data Wrangling 

I utilized the vegetation metrics dataset collected as part of the Pacific Northwest Forests project. Before merging, all metrics were cleaned utilizing a preparation script generated by the Ponisio lab. I also used published Differenced Normalized Burn Ratio (dNBR) data [@usdaforestserviceArcGIS2022] for map creation. A supplemental cleaning step was added by me to the vegetation cleaning script in which data was compared to csv of all ORFlora data from the Holiday Farm Fire area to confirm species presence and likeliness. Vegetation counts were standardized across umbel species due to collector miscounts. Shown in the cleaning script below, vegetation metrics were filtered to only relevant sites within the Holiday Farm Fire, and all burnpile plots (seeded enhancements and controls) were dropped from analysis. From here, a vegetation matrix was created for use in analysis. All sites were categorized by unique Stand, Row, Year combination labeled as "SiteID" and the sum of every unique PlantGenusSpecies was generated from number of plants. Due to species richness and Shannon not calculating correctly in the original cleaning script; I regenerated these numbers for every unique SiteID combination to use in further analysis. 

```{r}
#Library set up 
library(dplyr)
library(lmerTest)
library(lme4)
library(ggplot2)
library(ggeffects)
library(car)
library(ggpubr)
library(DHARMa)
library(terra)
library(stringr)
library(vegan)
library(tidyr)
library(tibble)
library(classInt)
```

```{r}
#Data Import 
load('/Users/laurenberger/Library/CloudStorage/Dropbox-UniversityofOregon/Lauren Berger/pnw_survey/data/veg_metrics_fire.Rdata')

```


```{r}
# ============================
# 2. Filter vegetation data
# ============================
outside.holiday <- c("181:Holiday_unburnt", "172:Johnson", "182:Ritchie",
                     "176:Holiday_unburnt", "174:Holiday_unburnt")

veg.filtered <- veg.metrics %>%
  filter(FireName == "Holiday",
         !grepl("6", SiteTransect),
         !grepl("6Mix|6Unenhanced", StandTrt),
         !Stand %in% outside.holiday) %>%
  mutate(SiteID = paste(Stand, Year, Round, sep = "_"))

# ============================
# 3. Creating a vegetation matrix, Summarize abundance per site × species for later use
# ============================
veg.mat <- veg.filtered %>%
  group_by(SiteID, PlantGenusSpecies) %>%
  summarise(Abundance = sum(NumPlant, na.rm = TRUE), .groups = "drop") %>%
  # Replace blank species names
  mutate(PlantGenusSpecies = ifelse(PlantGenusSpecies == "", "Unknown", PlantGenusSpecies)) %>%
  pivot_wider(names_from = PlantGenusSpecies, values_from = Abundance, values_fill = 0) %>%
  column_to_rownames("SiteID") %>%
  as.matrix()

# ============================
# 4. Calculate alpha diversity metrics
# ============================
alpha_div <- veg.filtered %>%
  group_by(SiteID) %>% 
  summarise(
    Richness = length(unique(PlantGenusSpecies)),  # simple species count
    Shannon = exp(vegan::diversity(NumPlant, index = "shannon")), # effective number of species
    .groups = "drop"
  )

# ============================
# 5. Add metadata
# ============================
site_info <- veg.filtered %>%
  distinct(SiteID, Stand, Year, Round, OwnerCode, Elevation, HighSeverityArea, DateStart, Long, Lat) %>%
  left_join(alpha_div, by = "SiteID")

# ============================
# 6. Check
# ============================
head(site_info)
summary(site_info$Richness)
summary(site_info$Shannon)
hist(site_info$Richness)
hist(site_info$Shannon)
```

## III. Analysis

For my analysis of alpha diversity, I generate a linear mixed model (completed in VegDiversity.r script) and  a Principal Coordinate Analysis (in VegPCoA.r script). I was able to combine both tests below, but renamed data frames for ease. The site_info data frame generated above contained elevation and HighSeverityArea (hereafter HSA) of each unique Stand, Year, Round combination. Both elevation and HSA were logged to improve model fit. The owner category “M” was used as the reference level for site-owner comparisons. I fit separate linear mixed-effects models for richness and Shannon diversity using the following formula: lmer(Richness or Shannon ~ OwnerCode + scale(log_Elevation) + scale(log_HighSeverityArea) + (1 | Round) + (1 | Stand) + (1 | Year), data = site_info). In these models, OwnerCode, log_Elevation, and log_HighSeverityArea were included as fixed effects, while Round, Stand, and Year were included as random intercepts to account for the hierarchical sampling design and repeated measurements across space and time.Model diagnostics performed using the DHARMa package indicated the presence of significant outliers and overall model misfit. Future analyses will re-examine these outliers and consider model refinements to improve fit. When evaluating species richness, HSA had no significant effect on richness (Estimate ≈ 0.03, p = 0.929), but increase in elevation showed a marginal negative trend (Estimate ≈ -1.13, p = 0.017). For Shannon diversity, HSA had no detectable effect on Shannon diversity  (Estimate ≈ 1.591, p = 0.27). Elevation had a significant negative effect (Estimate ≈ -5.242, p = 0.00599).
 

To complete my PCoA, I first separated the HSA surrounding each Stand into 5 separate categories. I used Jenks natural breaks classification from the classInt package to find natural groupings within the data. To prepare for permanova analysis, I further cleaned site_info as errors occurred from missing Elevation in three stands. I utilized vegdist from the vegan package to calculate Bray-Curtis dissimilarity between Stands and preformed a permanova test using adonis to determine if community composition differs between groups and along a continuous gradient. HSA explained only a small portion of variation in plant community composition, with the continuous severity measure accounting for 0.8% (R² = 0.008, F = 4.46, p = 0.001) and categorical severity explaining 2.1% (R² = 0.021, F = 2.81, p = 0.001). Next, I preformed a permutation test for homogeneity of dispersion from group centroids, indicating no significant differences among the five fire severity categories (F = 1.72, p = 0.144), suggesting that that sites within each fire severity category exhibit similar levels of beta‐diversity dispersion. I also completed a pairwise comparisons, with results of  Low severity group differed marginally from High (p = 0.041) and Very High (p = 0.020) categories, but these patterns are weak given the nonsignificant overall tests (F = 1.72, p = 0.127). 

```{r}
# ============================
# Log Variables and Preform Richness and Shannon LMER
# ============================
# ----------------------------
# Log-transform site variables
# ----------------------------
site_info <- site_info %>%
  mutate(
    log_HighSeverityArea = log(HighSeverityArea),
    log_Elevation = log(Elevation)
  )

# Make OwnerCode a factor and set baseline if needed
site_info$OwnerCode <- relevel(factor(site_info$OwnerCode), ref = "M")

# ----------------------------
# Plant Richness model
# ----------------------------
richness_model <- lmer(
  Richness ~ OwnerCode + scale(log_Elevation) + scale(log_HighSeverityArea) + (1 | Round) +  (1 | Stand) + (1 | Year),
  data = site_info
)

summary(richness_model)
```


```{r}
# Residual diagnostics
sim_res_rich <- simulateResiduals(richness_model)
plot(sim_res_rich)
testDispersion(sim_res_rich)
testZeroInflation(sim_res_rich)

# Extract effect for plotting
richness_sum <- summary(richness_model)
p_rich <- richness_sum$coefficients["scale(log_HighSeverityArea)", "Pr(>|t|)"]
slope_rich <- richness_sum$coefficients["scale(log_HighSeverityArea)", "Estimate"]
t_rich <- richness_sum$coefficients["scale(log_HighSeverityArea)", "t value"]

richness_eff <- ggpredict(richness_model, terms = "log_HighSeverityArea")
```


```{r}
# ----------------------------
# Shannon Diversity model
# ----------------------------
shannon_model <- lmer(
  Shannon ~ OwnerCode + scale(log_Elevation) + scale(log_HighSeverityArea) + (1 | Round) + (1 | Stand) + (1 | Year),
  data = site_info
)

summary(shannon_model)
```

```{r}
# Residual diagnostics
sim_res_shan <- simulateResiduals(shannon_model)
plot(sim_res_shan)
testDispersion(sim_res_shan)
testZeroInflation(sim_res_shan)

# Extract effect for plotting
shannon_sum <- summary(shannon_model)
p_shan <- shannon_sum$coefficients["scale(log_HighSeverityArea)", "Pr(>|t|)"]
slope_shan <- shannon_sum$coefficients["scale(log_HighSeverityArea)", "Estimate"]
t_shan <- shannon_sum$coefficients["scale(log_HighSeverityArea)", "t value"]

shannon_eff <- ggpredict(shannon_model, terms = "log_HighSeverityArea")


```

```{r}
# ----------------------------
# PCoA Calculations
# ----------------------------

# ----------------------------
# Categorize HighSeverity using "Jenks"
# ----------------------------
breaks <- classIntervals(site_info$HighSeverityArea, n = 5, style = "jenks")$brks

site_info <- site_info %>%
  mutate(
    HighSeverityCat = cut(HighSeverityArea, breaks = breaks, include.lowest = TRUE,
                          labels = c("Very Low", "Low", "Medium", "High", "Very High"))
  )

# ----------------------------
# . PERMANOVA prep
# ----------------------------
# Keep only complete cases, removing some NAs that exist in elevation and throw off model
site_info_clean <- site_info %>%
  filter(!is.na(HighSeverityArea),
         !is.na(Elevation),
         !is.na(OwnerCode)) %>%
  distinct(SiteID, .keep_all = TRUE)

# Keep only sites in veg.mat
common_sites <- intersect(rownames(veg.mat), site_info_clean$SiteID)
veg.mat.bray <- veg.mat[common_sites, ]
site_info_clean <- site_info_clean[site_info_clean$SiteID %in% common_sites, ]

# Reorder metadata to match veg.mat
site_info_clean <- site_info_clean[match(rownames(veg.mat.bray), site_info_clean$SiteID), ]
stopifnot(all(rownames(veg.mat.bray) == site_info_clean$SiteID))

# ----------------------------
# . Bray and PERMANOVA 
# ----------------------------
#creating dissimilarity values using Bray-Curtis method
dist_bray <- vegdist(veg.mat.bray, method = "bray")

#calculating permanova
set.seed(123)
permanova_res <- adonis2(veg.mat.bray ~ HighSeverityCat, data = site_info_clean, method = "bray", permutations = 999)
permanova_cont <- adonis2(veg.mat.bray ~ HighSeverityArea, data = site_info_clean, method = "bray", permutations = 999)

permanova_res
permanova_cont
```

```{r}
# ----------------------------
# Beta-dispersion visualization 
# ----------------------------
disp <- betadisper(dist_bray, site_info_clean$HighSeverityCat)
anova(disp)
permutest(disp, 999)

disp_df <- data.frame(
  Distances = disp$distances,
  FireSeverityCat = disp$group
)

```

## IV.Visualizations


```{r}
# ----------------------------
#Richness
# ----------------------------
ggplot(richness_eff, aes(x = x, y = predicted)) + 
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightblue", alpha = 0.3) +
  geom_line(color = "darkblue", size = 1.2, linetype = "dashed") +
  labs(
    x = paste0("Log High Severity Area\nSlope = ", signif(slope_rich, 3),
               "  t = ", signif(t_rich, 3),
               "  p = ", signif(p_rich, 3)),
    y = "Predicted Plant Richness",
    title = "Effect of High Severity Fire on Plant Richness"
  ) +
  theme_minimal(base_size = 14)
```
Figure 1. Effect of high-severity fire area on predicted plant richness. Predicted plant richness remains almost constant across the range of log-transformed high-severity area. The fitted regression line (dark blue) shows no meaningful trend, and the 95% confidence band (light blue) indicates high uncertainty relative to the near-zero effect size. 

```{r}
# ---- Shannon ----
ggplot(shannon_eff, aes(x = x, y = predicted)) + 
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightgreen", alpha = 0.3) +
  geom_line(color = "darkgreen", size = 1.2, linetype = "dashed") +
  labs(
    x = paste0("Log High Severity Area\nSlope = ", signif(slope_shan, 3),
               "  t = ", signif(t_shan, 3),
               "  p = ", signif(p_shan, 3)),
    y = "Predicted Shannon Diversity",
    title = "Effect of High Severity Fire on Shannon Diversity"
  ) +
  theme_minimal(base_size = 14)
```
Figure 2. Effect of high-severity fire area on predicted Shannon diversity. Predicted Shannon diversity shows a slight positive trend with increasing log-transformed high-severity area. However, the effect is not statistically significant. The fitted line (dark green) and its 95% confidence band (light green) illustrate substantial uncertainty around this relationship, indicating that high-severity fire area does not meaningfully influence Shannon diversity.



```{r}
# ---- Boxplots by Year ----
ggplot(site_info, aes(x = factor(Year), y = Richness, fill = factor(Year))) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  labs(x = "Year", y = "Plant Richness", title = "Plant Richness by Year") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none", plot.title = element_text(face = "bold", hjust = 0.5))
```
Figure 3A. Plant richness across survey years.
Boxplots show the distribution of plant richness from 2021 through 2025. Richness varies among years, with 2023 exhibiting the lowest median values and 2025 showing the highest median and broadest range. These differences may be due to sampling efforts, or differences species richness during post-fire recovery.

```{r}
ggplot(site_info, aes(x = factor(Year), y = Shannon, fill = factor(Year))) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  labs(x = "Year", y = "Shannon Diversity Index", title = "Shannon Diversity by Year") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none", plot.title = element_text(face = "bold", hjust = 0.5))

```
Figure 3B.Boxplots illustrate Shannon diversity values across the same period. Similar to richness, diversity shows year-to-year variation, with 2023 having the lowest medians and 2025 displaying both higher median values and greater overall variability. Again, these differences may be due to sampling efforts, or differences community composition during post-fire recovery.


```{r}
# ----------------------------
#Beta Diversity Visuals
# ----------------------------

pcoa_res <- cmdscale(dist_bray, eig = TRUE, k = 2)

var_PC1 <- pcoa_res$eig[1] / sum(pcoa_res$eig)
var_PC2 <- pcoa_res$eig[2] / sum(pcoa_res$eig)

plot_df <- data.frame(
  SiteID = rownames(veg.mat.bray),
  PC1 = pcoa_res$points[,1],
  PC2 = pcoa_res$points[,2]
) %>% left_join(site_info, by = "SiteID")

# ----------------------------
# PCoA plot
# ----------------------------
ggplot(plot_df, aes(x = PC1, y = PC2, color = HighSeverityCat, fill = HighSeverityCat)) +
  # points
  geom_point(size = 2.5, alpha = 0.7,
             position = position_jitter(width = 0.02, height = 0.02)) +
  # smaller, lightly filled ellipses
  stat_ellipse(
    geom = "polygon",
    alpha = 0.2,       # light fill
    level = 0.75,      # smaller ellipse
    color = NA         # no outline
  ) +
  scale_color_viridis_d(option = "plasma") +
  scale_fill_viridis_d(option = "plasma") +
  xlab(paste0("PCoA 1 (", signif(var_PC1*100,1), "%)")) +
  ylab(paste0("PCoA 2 (", signif(var_PC2*100,1), "%)")) +
  labs(title = "PCoA of Plant Communities by Fire Severity") +
  
  theme_minimal(base_size = 14)
```
Figure 4. Principal Coordinates Analysis (PCoA) of plant community composition across fire severity categories.
PCoA based on Bray–Curtis dissimilarity shows substantial overlap in community composition among the five fire‐severity categories (Very Low to Very High). The first two axes explain 10% and 9% of the variation, respectively. Although centroids and ellipses differ slightly across groups, large overlap indicates that plant communities are broadly similar across severity classes, consistent with the weak but statistically significant PERMANOVA results.

```{r}
# ----------------------------
# Beta-dispersion boxplot by Fire Category
# ----------------------------
ggplot(disp_df, aes(x = FireSeverityCat, y = Distances, fill = FireSeverityCat)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "Fire Severity Category",
       y = "Distance to Group Centroid",
       title = "Beta Dispersion by Fire Severity Category") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

```
Figure 5. Multivariate dispersion (beta dispersion) by fire severity category.
Boxplots show distances from each stand to its group centroid, representing within-category beta dispersion. Dispersion was not significantly different among the five fire‐severity categories. Although a few pairwise comparisons (e.g., Low vs. High; Low vs. Very High) showed marginally low p-values, these contrasts are not supported by the nonsignificant overall test.

```{r}
# ============================
# Mapping
# ============================
fire.rast <- rast('/Users/laurenberger/Library/CloudStorage/Dropbox-UniversityofOregon/Lauren Berger/pnw_survey_saved/spatial_data/fire_severity/HolidayFarm/or4417212223120200908_20200904_20200929_dnbr_utm.tif')
burn_boundary <- vect('/Users/laurenberger/Library/CloudStorage/Dropbox-UniversityofOregon/Lauren Berger/pnw_survey_saved/spatial_data/fire_severity/HolidayFarm/or4417212223120200908_20200904_20200929_burn_bndy_utm.shp')

fire_masked <- mask(fire.rast, burn_boundary)

Holiday_rast_df <- as.data.frame(fire_masked, xy = TRUE)
colnames(Holiday_rast_df)[3] <- "Holiday_Severity"

coords <- as.matrix(site_info[, c("Long", "Lat")])
points_vect <- vect(coords, crs = "EPSG:4326")
points_proj <- project(points_vect, crs(fire.rast))
points_df <- as.data.frame(points_proj, geom = "XY")
points_df <- cbind(points_df, site_info)


ggplot() +
  geom_tile(data = Holiday_rast_df, aes(x = x, y = y, fill = Holiday_Severity), na.rm = TRUE) +
  geom_point(data = points_df, aes(x = x, y = y, color = Richness), size = 2, alpha = 0.9) +
  scale_fill_gradientn(colors = c("forestgreen", "yellow", "orange", "red"), name = "Fire Severity") +
  scale_color_gradient(low = "skyblue", high = "skyblue4") +
  coord_equal() +
  theme_void() +
  labs(title = "Holiday Fire Severity Map") +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())



```
Figure 6. Spatial distribution of fire severity and plant richness across the Holiday Fire footprint.
The map displays fire severity across the Holiday Fire landscape taken fomr [@usdaforestserviceArcGIS2022], with severity values derived from remotely sensed burn-severity indices. Circles represent sampling locations colored by observed plant richness.


```{r}

# ============================
# Stand Age
# ============================

# # List layers in the geodatabase
# st_layers("/Users/laurenberger/Library/CloudStorage/Dropbox-UniversityofOregon/Lauren Berger/pnw_survey_saved/spatial_data/DataExport_1km.gdb")
# 
# #read in stand age layer
# stand_age <- layer <- st_read("/Users/laurenberger/Library/CloudStorage/Dropbox-UniversityofOregon/Lauren Berger/pnw_survey_saved/spatial_data/DataExport_1km.gdb", layer = "Stands_1km")
# 
# #read in stands
# stands <- read.csv('/Users/laurenberger/Library/CloudStorage/Dropbox-UniversityofOregon/Lauren Berger/pnw_survey_saved/data/raw/stands.csv')
# 
# #filter for Holiday only
# stands <- stands %>%
#   filter(grepl("holiday", StandCode, ignore.case = TRUE), 
#          OwnerCode == 'M')
# 
# #transform for multipolygon
# stand_age_clean <- st_cast(stand_age, "MULTIPOLYGON")
# 
# #remove any NAs (shouldn't exist)
# stand_points_clean <- stands %>%
#   filter(!is.na(Long) & !is.na(Lat))
# 
# #match up CRS
# stand_points_sf <- st_as_sf(stand_points_clean,
#                          coords = c("Long", "Lat"),
#                          crs = st_crs(stand_age))
# #assign 4326
# st_crs(stand_points_sf) <- 4326
# 
# #transform
# stand_points_sf <- st_transform(stand_points_sf, st_crs(stand_age))
# 
# #join!
# locations_with_age <- st_join(stand_points_sf,
#                               stand_age_clean[, "Age"],
#                               join = st_intersects)
# 
# #view
# head(locations_with_age)
# table(locations_with_age$Age)
# 
# #reviewing NAs
# locations_na <- locations_with_age %>%
#   filter(is.na(Age))
# 
# #viewing NAs
# ggplot() +
#   geom_sf(data = stand_age_clean, fill = "lightgrey", color = "black") +
#   geom_sf(data = locations_na, aes(color = is.na(Age)), size = 2) +
#   theme_minimal()
# 
# #trying to fill NAs by adjusting meters
# stand_age_buffer <- st_buffer(stand_age_clean, dist = 30)  # adjust meters
# locations_with_age <- st_join(stand_points_sf, 
#                               stand_age_buffer[, "Age"], 
#                               join = st_intersects)
# 
# ggplot() +
#   geom_sf(data = stand_age_clean, fill = "lightgrey", color = "black") +
#   geom_sf(data = locations_with_age, aes(color = is.na(Age)), size = 2) +
#   theme_minimal()

```

## V. Conclusion and Future Goals 

Alpha diversity analyses showed no significant relationship between the amount of high-severity area (HSA) surrounding stands and either species richness or Shannon diversity. Within sampled stands, high-severity fire did not significantly reduce local plant diversity. Elevation explained more variation in Shannon diversity than HSA, although most of the variation in both richness and Shannon diversity was attributable to site-specific factors captured by the random effects (Stand, Year, Round).

Beta diversity analysis using PCoA and PERMANOVA revealed significant differences in group centroids among fire-severity categories; however, these differences accounted for only a small proportion of total community variation. Tests of multivariate dispersion (ANOVA and permutest) indicated no significant differences in within-group dispersion among severity categories. Although a few pairwise comparisons (Low vs. High; Low vs. Very High) showed marginal differences, these results should be interpreted cautiously given the nonsignificant overall dispersion test. Collectively, these findings suggest limited evidence for differences in turnover within severity groups, though modest shifts in community composition occur between groups.

Future analyses should incorporate additional environmental variables that may better explain variation in community structure. Stand age prior to fire and growing degree days are likely important predictors and will be included in the next phase of modeling. Considering functional traits may also reveal meaningful ecological patterns not captured by richness or Shannon diversity alone. Previous studies have reported mixed effects of fire severity on plant diversity, with some documenting increased richness at high-severity sites [@brodieProductivityModifiesEffects2021] and others reporting declines [@strandDoesBurnSeverity2019]. Both studies, however, found strong differences in functional traits of species occupying high-severity areas, particularly an increase in species associated with residual or colonizing strategies. These patterns may help explain stand-specific variation observed in this dataset.


## VI. References
